---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# PredImption: Mutation Impact Predition Based on Network Properties

<!-- badges: start -->
<!-- badges: end -->

# Overview

'PredImption' is a tool for impact prediction of a mutation on a protein structure by using network properties. 

# Installation

You can install the released version of PredImption from [CRAN](https://CRAN.R-project.org) with:

``` {r installation_guide1, eval=FALSE}
install.packages("PredImption")
```

or install via 'devtools':

``` {r installation_guide2, eval=FALSE}
library(devtools)
devtools::install_github("berkgurdamar/PredImption")
```

# Usage

'PredImption' uses data.frame structures as an input. data.frame should consist "PDB_ID", "Chain", "Position", "Orig_AA", "Mut_AA":

```{r example_input, echo=FALSE}
test_data <- as.data.frame(rbind(c("2DN2", "B", 1, "VAL", "ALA"),
                                 c("2DN2", "B", 6, "GLU", "ALA")))
colnames(test_data) <- c("PDB_ID", "Chain", "Position", "Orig_AA", "Mut_AA")
knitr::kable(test_data, align = c("c", "c", "c", "c", "c"))
```

Mutation impact prediction can be done via `PredImption()` function:

`PredImption()` function works on each PDB ID respectively. First, downloads the PDB file, creates distance matrix and turns PDB structure into a network. Calculates _Eigen Centrality Z-Score, Shortest Path Z-score_ and _Betweenness Z-Score_ of input positions. Find the _Mutation Type (1-Deadly, 0-Not Deadly)_, find the _Gene Name_ from the PDB file and if there are multiple genes annotated for the input PDB file, ask user to choose the input gene, gets _GnomAD Synonymous Z-Score, Non-Synonymous Z-Score_ and _PLoF Score, Genic Intolerance Score_, gets _BLOSUM62 score_ of the amino acid change, finds the _KEGG Pathway Number_ which contains the input gene. Finally, make prediction based on an _Adaboost_ model and classifies the mutation as _Disease Causing_ or _Silent_.


```{r usage, eval=FALSE}
library(PredImption)
pred_res <- PredImption(input_df)
```

`PredImption()` function returns a data.frame which contains additional two columns; _Prediction_ and _Probability_. _Prediction_ represents the result of the impact prediction and _Probability_ represents the probability that the mutation classified as _Disease Causing_ or _Silent_.

```{r example_output, echo=FALSE}
example_result <- as.data.frame(rbind(c("2DN2", "B", 1, "VAL", "ALA", "silent", "0.7676430"),
                                      c("2DN2", "B", 6, "GLU", "ALA", "silent", "0.7225815")))

colnames(example_result) <- c("PDB_ID", "Chain", "Position", "Orig_AA", "Mut_AA", "Prediction", "Probability")

knitr::kable(example_result, align = c("c", "c", "c", "c", "c", "c", "c"))
```

